### 2016-10-26 sec-results.tex
Add missing figures and latent state transition results.

\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 1} & \textbf{State 2} & \textbf{State 3} & \textbf{State
    4}\\\hline
    \includegraphics[width=0.22\textwidth]{figures/text-2state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-2state/state1.png}\\
    \includegraphics[width=0.22\textwidth]{figures/text-3state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-3state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-3state/state2.png}\\
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state2.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state3.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\textretrieval{} MOOC.} % I don't know why \protect is needed
  \label{fig:text-state-evolution}
\end{figure*}

In Figure~\ref{fig:text-state-evolution} we see the evolution of the latent
state representations found for the \textretrieval{} MOOC\footnote{Due to
space constraints, the graphs are quite small, but the states are in the
same positions as they were in previous figures to aid interpretation.}.
With $K=2$ we uncover a video watching pattern (state 2) and a course
material browsing pattern (state 1). However, we can see that when $K=3$ we
are able to uncover a new pattern involving forum behavior in state 3
(notice the node sizes on the bottom right). As we increase $K$ to four, we
can see that state 2 splits into state 2 and state 4. These states appear
quite similar at a glance, but there are still some key differences. First,
we can see that state 2 now has a non-negligible forum component, whereas
state 4 has hardly any weight on the forum.

\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 1} & \textbf{State 2} & \textbf{State 3} & \textbf{State
    4}\\\hline
    \includegraphics[width=0.22\textwidth]{figures/sustain-2state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-2state/state1.png}\\
    \includegraphics[width=0.22\textwidth]{figures/sustain-3state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-3state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-3state/state2.png}\\
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state2.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state3.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\sustain{} MOOC.} % I don't know why \protect is needed
  \label{fig:sustain-state-evolution}
\end{figure*}

We observe similar behavior in Figure~\ref{fig:sustain-state-evolution}
as we increase $K$ when fitting our TL-HMM on the \sustain{} MOOC. Again,
in the transition between $K=2$ and $K=3$ we discover forum behavior
patterns as a latent state. However, in the transition between $K=3$ to
$K=4$ we instead see a refining of that discovered forum state, where
state 3 captures asymmetric transition probabilities between ``forum thread
view'' and ``forum thread list''. These states could be seen as redundant,
in which case a more appropriate setting of $K=3$ may be appropriate for
the \sustain{} dataset.

\subsection{Transitions Between Latent States}
% using k=6, show that we can uncover interesting patterns in how students
% transition between the different latent states; argue that this discovery
% is interesting and unique to our formulation and could not be discovered
% before

>>>>

\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 0} & \textbf{State 1} & \textbf{State 2} & \textbf{State
    3}\\\hline
    \includegraphics[width=0.22\textwidth]{figures/text-2state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-2state/state1.png}\\
    \includegraphics[width=0.22\textwidth]{figures/text-3state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-3state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-3state/state2.png}\\
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state2.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/text-4state/state3.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\textretrieval{} MOOC.} % I don't know why \protect is needed
  \label{fig:text-state-evolution}
\end{figure*}

In Figure~\ref{fig:text-state-evolution} we see the evolution of the latent
state representations found for the \textretrieval{} MOOC\footnote{Due to
space constraints, the graphs are quite small, but the states are in the
same positions as they were in previous figures to aid interpretation.}.
With $K=2$ we uncover a video watching pattern (state 1) and a course
material browsing pattern (state 0). However, we can see that when $K=3$ we
are able to uncover a new pattern involving forum behavior in state 3
(notice the node sizes on the bottom right). As we increase $K$ to four, we
can see that state 1 splits into state 1 and state 3. These states appear
quite similar at a glance, but there are still some key differences. First,
we can see that state 1 now has a non-negligible forum component, whereas
state 3 has hardly any weight on the forum.

\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 0} & \textbf{State 1} & \textbf{State 2} & \textbf{State
    3}\\\hline
    \includegraphics[width=0.22\textwidth]{figures/sustain-2state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-2state/state1.png}\\
    \includegraphics[width=0.22\textwidth]{figures/sustain-3state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-3state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-3state/state2.png}\\
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state0.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state1.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state2.png}
    &
    \includegraphics[width=0.22\textwidth]{figures/sustain-4state/state3.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\sustain{} MOOC.} % I don't know why \protect is needed
  \label{fig:sustain-state-evolution}
\end{figure*}

We observe similar behavior in Figure~\ref{fig:sustain-state-evolution}
as we increase $K$ when fitting our TL-HMM on the \sustain{} MOOC. Again,
in the transition between $K=2$ and $K=3$ we discover forum behavior
patterns as a latent state. However, in the transition between $K=3$ to
$K=4$ we instead see a refining of that discovered forum state, where
state 2 captures asymmetric transition probabilities between ``forum thread
view'' and ``forum thread list''. These states could be seen as redundant,
in which case a more appropriate setting of $K=3$ may be appropriate for
the \sustain{} dataset.

\subsection{Transitions Between Latent States}

\begin{figure*}
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth,trim={0 3cm 0 2cm}]{figures/trans-comp/trans-avg.png}
    \caption{\label{fig:trans-avg}}
  \end{subfigure}%
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth,trim={0, 3cm 0 2cm}]{figures/trans-comp/trans-perfect.png}
    \caption{\label{fig:trans-perfect}}
  \end{subfigure}
  \caption{The latent state transition diagrams for a 4-state TL-HMM fit to
  \protect\textretrieval{} for all students (a) compared to only ``perfect''
  students (b).}
  \label{fig:trans-comp}
\end{figure*}

A unique property of our model is its ability to capture transitions
between the \emph{behavior patterns themselves} that are captured by the
latent states. In Figure~\ref{fig:trans-avg} we show the latent state
transition diagram for a 3-state TL-HMM fit on \textretrieval{}. We can
immediately observe two things: \begin{enumerate*}[label=(\arabic*)]
  \item each latent state has a very high ``staying'' probability, and
  \item the prevalence of each latent state matches our intuition.
\end{enumerate*}
In particular, we can see that the forum browsing state (state 2) has
relatively lower probability than the other states as we might expect. It
also makes sense that state 0 has rather high probability as this state
likely captures all sequences where a student logged in to the platform and
then did nothing else (likely checking for updates). If we look at state 1
and state 3, their relative probabilities match our intuition as
well. State 1 seems to capture a more engaged browsing session, where there
is non-negligible probability associated with different activities such as
quiz taking and forum browsing and, importantly, these activities have high
probability symmetric edges (so students are taking quizzes one after the
other, or viewing forum threads in succession). By contrast, state 3 seems
to capture a more passive student, with negligible probability mass
associated with forum activity (with low symmetry in the edges). The link
between ``quiz submit'' and ``quiz start'' (indicating quiz repetition) is
also significantly lower than state 1.

Thus, we might expect to see students that perform well in the course
preferring states 1 and 2 over states 0 and 3. To verify this, we took the
model we learned on the full training data and retrofit it to training data
consisting only of sequences produced by students in \textretrieval{} that
had perfect marks. To prevent the latent state meanings from drifting, we
forced the model parameters associated with their Markov model
representations to be fixed, in effect only learning initial and transition
probabilities for the top layer of our TL-HMM.\ We show the updated latent
state transition diagram in Figure~\ref{fig:trans-perfect}. While the
differences are small, we do in fact see what we suspected: states 1 and 2
have higher probability in the ``perfect'' student group compared to the
model learned on all students, and states 0 and 3 have lower probability.
